### 2017-11-26 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #23
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresss #36
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

ParSplice~\cite{perez:jctc20150parsplice} is an accelerated molecular dynamics
(MD) simulation package developed at LANL. It is part of the Exascale Computing
Project\footnote{http://www.exascale.org/bdec/} and is important to LANL's
Materials for the Future initiative. As shown in Figure~\ref{fig:parsplice},
the phases are:

@@

Changing growth rates and temperature alters the size, shape, and locality of
the data store keyspace. Lower temperatures and smaller growth rates create
hotter keys with smaller keyspaces as many segments are generated in the same
set of states before the trajectory can escape to a new region of state space.
In Figure~\ref{fig:motivation-regimes}, the \(\Delta_1\) growth rate adds atoms
every 100K microseconds while the \(\Delta_2\) growth rate adds atoms every 1
million microseconds. So \(\Delta_2\) has a smaller growth rate resulting in
hotter keys (line on \(y1\) axis) and a smaller keyspace (dots on the \(y2\)
axis).  Values smaller than \(\Delta_2\)'s growth rate or a temperature of 400
degrees results in very little database activity because state transitions take
too long. Similarly, values larger than \(\Delta_1\)'s growth rate or a
temperature of 4000 degrees result in an equally meaningless simulation as
transitions are unrealistic.

@@

\subsection{ParSplice Keyspace Analysis}
\label{sec:parsplice-keyspace-analysis}
%IT HAS 8K keys!  How did we take these measurements

@@

\subsubsection*{An active but small keyspace}
Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
pairs from the worker's cache for two different initial conditions of
\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
the read request rate (\(y1\) axis) and the dots along the bottom are the
number of unique keys accessed (\(y2\) axis).  This figure demonstrates that
small changes to \(\Delta\) can have a strong effect on the timing and
frequency with which new EOM minima are discovered and referenced.  The
ParSplice caches on each node are trimmed when memory pressure reaches a
threshold but the number of unique keys accessed in
Figure~\ref{fig:motivation-regimes} suggests that a small cache of EOM minima
should be sufficient.

@@

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run (this motivates Section~\S\ref{sec:dom-specific}).
The growth rate, temperature, and number of workers have a predictable effect
on the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
the number of reads changes with different initial conditions (\(\Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).
Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
rates has temporal locality, as the reads per second for \(\Delta_2\) looks
like the reads per second for \(\Delta_1\) stretched out along the time axis.
Trends also exist for temperature and number of workers but are omitted here
for space. This structure means that we can learn the key access patterns and adapt the
storage system (this motivates Section~\S\ref{sec:dom-specific}).

>>>>

ParSplice~\cite{perez:jctc20150parsplice} is an accelerated molecular dynamics
(MD) simulation package developed at LANL. It is part of the Exascale Computing
Project\footnote{http://www.exascale.org/bdec/} and is important to LANL's
Materials for the Future initiative.

\subsection{Background}

As shown in Figure~\ref{fig:parsplice}, the phases are:

@@

Changing growth rates and temperature alters the size, shape, and locality of
the keyspace. Lower temperatures and smaller growth rates create
hotter keys with smaller keyspaces as many segments are generated in the same
set of states before the trajectory can escape to a new region of state space.
In Figure~\ref{fig:motivation-regimes}, the \(\Delta_1\) growth rate adds atoms
every 100K microseconds while the \(\Delta_2\) growth rate adds atoms every 1
million microseconds. So \(\Delta_2\) has a smaller growth rate resulting in
hotter keys (line on \(y1\) axis) and a smaller keyspace (dots on the \(y2\)
axis).  Values smaller than \(\Delta_2\)'s growth rate or a temperature of 400
degrees results in very little database activity because state transitions take
too long. Similarly, values larger than \(\Delta_1\)'s growth rate or a
temperature of 4000 degrees result in an equally meaningless simulation as
transitions are unrealistic.

@@

\subsection{Results and Conclusions}
\label{sec:parsplice-keyspace-analysis}
%IT HAS 8K keys!  How did we take these measurements

@@

\subsubsection*{An active but small keyspace}
Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
pairs from the worker's cache for two different initial conditions of
\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
the read request rate (\(y1\) axis) and the dots along the bottom are the
number of unique keys accessed (\(y2\) axis).  This figure demonstrates that
small changes to \(\Delta\) can have a strong effect on the timing and
frequency with which new EOM minima are discovered and referenced. This finding
suggests that we need a flexible policy language and engine.

@@

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run (this motivates Section~\S\ref{sec:dom-specific}).
The growth rate, temperature, and number of workers have a predictable effect
on the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
the number of reads changes with different initial conditions (\(\Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).
Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
rates has temporal locality, as the reads per second for \(\Delta_2\) looks
like the reads per second for \(\Delta_1\) stretched out along the time axis.
Trends also exist for temperature and number of workers but are omitted here
for space. This structure means that we can learn the key access patterns and adapt the
storage system accordingly (this motivates Section~\S\ref{sec:dom-specific}).
